---
title: "Colorblind fridenly plotting in R"
author: "Jie Hua"
date: "2025-08-28"
date-modified: "2025-08-28"
categories: [Plot]
img: "index.png"
---

## 0. Basic Information

To use colorblind-friendly palettes in **R**, you mainly need the following packages. Be sure that we checked the plots with these packages before saving the plots. Thank you!

## 1. colorbindr

> Simulate colorblindness in production-ready R figures. Written by Claire D. McWhite and Claus O. Wilke.

Official repo: <https://github.com/clauswilke/colorblindr>

### 1.1 Installation

```{r}
#| warning: false
#| eval: false
remotes::install_github("wilkelab/cowplot")
install.packages("colorspace", repos = "http://R-Forge.R-project.org")
remotes::install_github("clauswilke/colorblindr")
```

It could also be installed with: (but you need to install [**pak**](https://github.com/r-lib/pak)first.

```{r}
#| eval: false
#| warning: false
pak::pkg_install("clauswilke/colorblindr")
```

### 1.2 Demo

It is easy to follow the official document, first make a ggplot object like:

```{r}
library(ggplot2)
fig <- ggplot(iris, aes(Sepal.Length, fill = Species)) + geom_density(alpha = 0.7)
fig
```

Then we could use this package to check in different simulations:

```{r}
library(colorblindr)
cvd_grid(fig)
```

It is also possible to inspect it in the interactive app, I just show the script, it will not open a tab for readers.

```{r}
#| eval: false
view_cvd(fig) # starts the interactive app
```

# 2. colorblindcheck

> Check Color Palettes for Problems with Color Vision Deficiency

Official repo: <https://github.com/Nowosad/colorblindcheck>

### 2.1 Installation

```{r}
#| warning: false
#| eval: false
install.packages("colorblindcheck")
```

### 2.2 Demo

```{r}
library(colorblindcheck)
rainbow_pal = rainbow(n = 7)
rainbow_pal
#> [1] "#FF0000" "#FFDB00" "#49FF00" "#00FF92" "#0092FF" "#4900FF" "#FF00DB"
palette_check(rainbow_pal, plot = TRUE)
```